---
title: "data_tidying"
author: "A Bokhobza"
date: "2023-03-09"
output: html_document
---

**Aim**: Tidy data from an excel file.\
**Contact**: [alexandre.bokhobza\@gmail.com](mailto:alexandre.bokhobza@gmail.com){.email}.\
**Reference used for creating this notebook**:[DataViz](https://joachimgoedhart.github.io/DataViz-protocols/read-and-reshape.html#tidy-data)

------------------------------------------------------------------------------

1. Load appropriate libraries.\

```{r warning=FALSE}

library(tidyverse)
library(openxlsx)
library(here)

```

1. Setup Here package for easy sharing between collaborators.\

```{r}

here::i_am("data_tidying.Rproj")
here()

```

1. Open and load the desired file as a dataframe (named df):\

```{r}

df <- read.xlsx("../data/analysis_Test.xlsx", fillMergedCells = TRUE, colNames = FALSE )

```
Adapt the file name according to the file to process.\

1. Create R objects (type = vector, named first_row ) from the column names of the excel file and replace the whitespace by a point\

```{r }

first_row <- df[1,] %>% unlist(use.names = FALSE)
first_row <- str_replace_all(first_row, pattern = " ", replacement = ".")

second_row <- df[2,] %>% unlist(use.names = FALSE)
second_row <- str_replace_all(second_row, pattern = " ", replacement = ".")

```  

1. Combine the vector (first_row and second_row) into a single one (use underscore during combination).\

```{r}

combined_labels <- paste(first_row, second_row, sep = "_")

```

1. Deletes the first rows from the dataframe.\

```{r }

df_no_header <- df[-c(1:2),]

```

1. Add the new lables to the df.\

```{r}

colnames(df_no_header) <- combined_labels

```

1. Tidy the data.\

```{r}

df_tidy <- pivot_longer(df_no_header, cols = everything(), names_to = "combined_labels", values_to = "Intensity")

```

1. Split the combined labels.\

```{r}

df_tidy_final <- df_tidy %>% separate(combined_labels, c('Strand', 'Target'), sep = "_")

```

1. Save the tidy df as a csv file.\

```{r}

write.csv(df_tidy_final, "../output/analysis_test_tidy.csv", row.names = FALSE)

```

